---
title: "MOEA/D"
subtitle: "MOEA/D as Cellular GA"
author: | 
  | Yuri Lavinas
  | Master Student - University of Tsukuba
date: "05/09/2018"

output: 
  beamer_presentation:
    slide_level: 3
    theme: "Szeged"
    fonttheme: "professionalfonts"
    citation_package: natbib
    latex_engine: "pdflatex"
    includes:
          in_header: header_pagenrs.tex
          after_body: after_body.tex
bibliography: bib.bib
link-citations: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(cache = TRUE)
library(png)
library(grid)
library(ggplot2)
library(gridExtra)
```

# MOP
## Multi-objective Problems
### What is MOP?
Multiobjective Optimization Problem have $m$ multiple objective functions that must be optimized simultaneously.

Minimize$^1$ $F(x) = (f_1(x), f_2(x), ..., f_m(x))$,

subject to $x$ in $\Omega$.

- $F(x)$ objective functions;
- $f_i$ is the i-th objective to be minimized;
- $x$ is the decision vector;
- $\Omega$ is the decision space.
 
 \footnotesize $^1$ All definitions are for minimization. Following inequalities should be reversed if the goal is to maximize.

### Why is MOP interesting?

1. Many real-world scientific and engineering are MOP.
    - Water quality control, Groundwater pollution remediation, Design of marine vehicles, ... @coello2007evolutionary.
    - Petrol extraction.
2. Hard problems: to balance the interests of the multi-objective as a whole is hard. 


## Pareto Set
### What is Pareto Set?

Objectives may be conflicting
    - The goal is to find good trade-offs.

- Set of solutions.

Set of *optimum solutions* - Pareto set.

- Non-dominated solutions: no single solution provides a better trade-off in all objectives.

### What is Pareto Set?

\centering
```{r fig.width=4.5, fig.height=10,echo=FALSE}
img <- readPNG("pareto_front_diff_scalarizins_f.png")
 grid.raster(img)
```

\tiny From @ishibuchi2009adaptation.


## Pareto Set
### Non-dominated solutions

1. Let $u = (u_1, ..., u_m)$ and $v = (v_1, ..., v_m)$ vectors in $\Omega$ (the decision space).
    - $\forall i:u$ dominates $v$ if $f_i(u) \leq f_i(v)$ and $\exists j:f_j(u) < f_j(v)$.
    - u dominates v, v is dominated by u, u is better that v.

2. A point $x^*$ in $\Omega$  is called *Pareto Optimal* if no other point dominates $x^*$. 
    <!-- - $\nexists F(y^*)$ that dominates $F(x^*)$ -->

## Pareto Set
### Pareto Front

1. The set of all Pareto Optimal is called the *Pareto Set*. 
    - $P^*$ = {$x \in \Omega:\nexists y \in \Omega$ and $F(y) \leq F(x)$}
3. **Pareto Front** is the image of the Pareto Set in the objective space.
    - PF = {$F(x) = (f_i(x), ..., f_m(x)): x \in P^*$}

### Pareto Front
\centering
```{r fig.width=5, fig.height=15,echo=FALSE}
img <- readPNG("pareto_dominated.png")
 grid.raster(img)
```
\tiny From: http://www.cenaero.be/Page.asp?docid=27103&langue=EN

# MOEA/D
## MOEA/D
### Decompostion

MOEA/D represents a class of population-based meta-heuristics for solving MultiObjective Problems (MOPs).

- It is based on decomposition - one kind of scalarizing function
  - One multi-objective problem becomes various single-objective sub-problems.
  - All sub-problems are solved in parallel. 
  - A decomposition strategy generates weight vectors that defines the sub-problems.

## MOEA/D
### Why use decomposition?

- It may be good at generating an even distribution of solutions in MOPs;
- It reduces the computation complexity when compared to other algorithms (NSGA-II) (at each generation), @zhang2009performance;
<!-- - An optimal solution of a set of scalar optimization problems can be a Pareto optimal solution, under mild conditions; -->
<!-- - All solutions can be compared based on their objective function values; -->
<!-- - It is simple to find a solution to multi single-objective problems than for a multi-objective problem; -->
- Fitness assignment and diversity maintenance become easier to handle.


## MOEA/D
### Components of the MOEA/D

- Decomposition strategy: decomposes w/ weight vectors;
- Aggregation function:  weight vector => single-objective sub-problems;
- **Neighborhood assignment strategy**: Relationship between sub-problems;
- Variation Stack: New candidates solutions;
- Update Strategy: Maintain/discard candidate solutions;
- Constraint handling method: Constraint violation;
- Termination Criteria: when to stop the search.

### Simple Visualization 

\centering
```{r fig.width=3.4, fig.height=10,echo=FALSE}
img <- readPNG("MOEAD.png")
 grid.raster(img)
```


# MOEA/D and Cellular GA

## Cellular GA
### What is Cellular GA?


- Cellular GA (cGA) is a distribuited model of the Genetic Algorithm (GA).
     - From  @gong2015distributed: A cGA has one population but arranges the individuals on the grid, called **cell**.

- Cellular GA is well explored in the context of single-objective problems (SOP).
    - Each individual is only compared with its neighborhood.
    - Each individual uses only its neighborhood for updates.
    - It is highly parallelized: Synchronous x Asynchronous.

### Cell and neighbors
\raggedleft
```{r fig.width=3, fig.height=8.5,echo=FALSE}
myimages <- list.files("images/", pattern = ".png", full.names = TRUE)
knitr::include_graphics(myimages)
```

\raggedleft \tiny From: @robles2009evolutionary


### Research focus

- Goal: To propose improvements to the MOEA/D by incorporating (well-)established techniques from the Cellular GA.

- Why?  MOEA/D can be viewed as a cGA.
    - A cell can be seen as a specific "Neighborhood assignment strategy", where each cell has its own weight vector.

### Research focus

\centering
```{r fig.width=4, fig.height=10,echo=FALSE}
img <- readPNG("all.png")
 grid.raster(img)
```



## Cellular GA and MOEA/D @ishibuchi2009adaptation
### MOEA/D as Cellular GA

1. Fitness of a solution is calculated at each cell using a scalarizing fitness function with the weight vector associated with that cell.
    - The cells are located in the (m-1)-dimensional subspace -> satisfying the normalization condition:
         - $w_1 + ... + w_m = 1$ and $w_i \geq 0$ for i = 1, 2 ,...,m.
    - Two neighborhood structures: 
         - Local selection.
         - Local replacement, both defined using the Euclidean distance between cells.

### MOEA/D as Cellular GA
2. Parent solutions are randomly selected from the **selection neighbors**.
    2. Offspring is generated from the parents by crossover, mutation and/or repair.
3. Generated offspring is compared with the **replacement neighbors**, considering the weight vector.

Obs: The local selection and local replacement are performed for the next cell after they are completed for the current cell. 

### Discussion

- The main characteristic feature of MOEA/D as a cGA is the use of **local replacement** in addition to local selection. 
    - Generated offspring for a cell is compared with not only the current solution of the cell but also its neighbors for possible replacement.
    
- Local replacement neighborhood has greater effect on the performance than local selection neighborhood.
    - Increasing its size tends to be better.

### Discussion

- Good results with very large populations.

- Multiple neighbors can be replaced by one good offspring - less diversity.
    - Decreasing the size of competition neighborhood increases the non-dominated solutions, but degrades the search ability of the MOEA/D.
    

# Interesting Recent Developments
## MOEA/D
### MOEA/D

1. Dual populations (external population).
    - Store the visited non-dominated solutions with a weight vector, and then add/remove sub-problems given crowded/sparse regions by adjusting the weight vector, @qi2014moea.
    - Stored the non-dominated solutions such that the Hyper Volume is maximized, @jiang2016towards.
    - External population is updated using non-dominates sorting and crowding distance <- NSGA-II. Also it is used to guide the allocation of computational resources to a sub-problem given its contribution, @cai2015external.

3. Multigrid scalarizing scheme or a single grid with different scalarizing functions are alternately assigned for each weight vector @ishibuchi2010simultaneous. 

## MOEA/D as Cellular GA
### MOEA/D as Cellular GA

1. Asynchronous - could degenerate the algorithm.
    - Dynamic computation resource allocation to different sub-problems, @zhang2009performance and, @zhou2016all.	
    
2. Island model with different parameter values and/or reproduction operators and/or scalarizing scheme - aiming to balance convergence and diversity, and to explore the objective space better.
    - Neighborhood relationship should be defined in the objective space and should be adaptive, @giagkiozis2014generalized.
    - (self-)Adaptive replacement strategies, @wang2016adaptive and @zhang2016self.


# References
## References
### Basic References
Most of this presentation was based on the works of @trivedi2017survey and  @zhang2007moea.

\small